The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-12 manuscript/manuscript.Rnw
fix fusciceps typo
cite amazing armadillo

\begin{table}
  \begin{tabular}{c c c p{3in}}
    \toprule
    Species                       & Gestation (days) & Litter Size  & Source \\
    \midrule
    \textit{Ateles fusciceps}     & 224 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Bos taurus}           & 278 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Canis familiaris}     & 61 & 4 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Dasypus novemcinctus} & 135 & 4 & \DLA{Cite Amazing Armadillo: geography book}\\
    \textit{Equus caballus}       & 336 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation}\\
    \textit{Loxodonta africana}   & 644 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Homo sapiens}         & 270 & 1.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Monodelphis domestica}& 15 & 7.1 & \citet{Harder.ea1993:Gestationandplacentationin} \\
    \textit{Mus musculus}         & 20 & 6.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Nannospalax galili}   & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals} \\
    \textit{Ovis aries}           & 151 & 1.35 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Pan paniscus}         & 223 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental} \\
    \textit{Spalax carmeli}       & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals}\\
    \textit{Sus scrofa}           & 114 & 5.8 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \bottomrule
  \end{tabular}
  \caption{Gestational periods in days and average liter sizes for studied species.}
  \label{tab:gestational_periods}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
print(xtable(top.spider.monkey.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.spider.monkey.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\smonkey}{tab:all_ateles_fuciceps_placenta}}
\label{tab:top50_spider_monkey_placenta}
\end{table}

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(\cref{tab:top50_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles fusciceps","PSG8")}), %
a member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
\cref{tab:top50_spider_monkey_placenta}), and multiple components of
the annexin complex %
(\Sexpr{gene.expression("ateles fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}. All genes with expression
greater than 50 \ac{FPKM} are listed in Supplementary
\cref{tab:all_ateles_fuciceps_placenta}.

>>>>

\begin{table}
  \begin{tabular}{c c c p{3in}}
    \toprule
    Species                       & Gestation (days) & Litter Size  & Source \\
    \midrule
    \textit{Ateles fusciceps}     & 224 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Bos taurus}           & 278 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Canis familiaris}     & 61 & 4 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Dasypus novemcinctus} & 150 & 4 & \citet{Smith1984:amazingarmadillo} \\
    \textit{Equus caballus}       & 336 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation}\\
    \textit{Loxodonta africana}   & 644 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Homo sapiens}         & 270 & 1.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Monodelphis domestica}& 15 & 7.1 & \citet{Harder.ea1993:Gestationandplacentationin} \\
    \textit{Mus musculus}         & 20 & 6.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Nannospalax galili}   & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals} \\
    \textit{Ovis aries}           & 151 & 1.35 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Pan paniscus}         & 223 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental} \\
    \textit{Spalax carmeli}       & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals}\\
    \textit{Sus scrofa}           & 114 & 5.8 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \bottomrule
  \end{tabular}
  \caption{Gestational periods in days and average liter sizes for studied species.}
  \label{tab:gestational_periods}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
print(xtable(top.spider.monkey.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.spider.monkey.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\smonkey}{tab:all_ateles_fusciceps_placenta}}
\label{tab:top50_spider_monkey_placenta}
\end{table}

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(\cref{tab:top50_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles_fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles_fusciceps","PSG8")}), %
a member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles_fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
\cref{tab:top50_spider_monkey_placenta}), and multiple components of
the annexin complex
(\Sexpr{gene.expression("ateles_fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}.
\Sexpr{gene.expression("ateles_fusciceps","AL365202.1",complete=TRUE)}
is also highly expressed, has no known function, and is duplicated
multiple times in \textit{Macaca mulatta}. All genes with expression
greater than 50 \ac{FPKM} are listed in Supplementary
\cref{tab:all_ateles_fusciceps_placenta}.
